ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-09 rebuttal.tex

\begin{reviewer}
    %
    16.~The Python package Cerberus is used to validate the format and content of the given ChemKED
    file. How does this package handles/reports the inconsistencies with the schemas? As far as my
    local test went with giving a corrupted YAML file to the ChemKED initiator, the package just
    spits out a Python error, which might be not the most user-friendly solution.
    %
\end{reviewer}

\begin{response}
    %
    We recognize that this is a limitation of the current version of PyKED. The technical reason for
    this in the Cerberus package has very recently been resolved (see
    \url{https://github.com/pyeve/cerberus/issues/278}), so we plan to update the error-reporting
    mechanism in PyKED to report well-formatted errors that have a detailed description of the
    error. We have opened a public issue to track the progress on this:
    \url{https://github.com/pr-omethe-us/PyKED/issues/78}. Thank you for suggesting this
    improvement.
    %
\end{response}

\begin{reviewer}
    %
    17.~A recurring issue with actively developing codes is backwards compatibility. How are the
    authors planning on addressing this issue?
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that backwards compatibility may become an issue. We have addressed
    this concern in ChemKED by including the database version as one of the required keys in the
    ChemKED file. Although the functionality doesn't exist yet, we intend to be able to check this
    version and report any feature deprecation or removal to the user (see
    \url{https://github.com/pr-omethe-us/PyKED/issues/22}), and we plan to provide an automated
    upgrade path for ChemKED files when backwards-incompatible changes are made. In addition, we use
    the widely supported Semantic Versioning paradigm, where changes in the various parts of the
    version number (vX.Y.Z) have specific meaning; changing the major version (X) indicates a
    backwards incompatible change, changing the minor version (Y) indicates somewhat large change,
    possibly requiring small backwards-compatibility fixes (but should generally be entirely
    backwards-compatible), and changing the patch version (Z) indicates a bug fix with no
    anticipated incompatibilities.

@@

\begin{response}
    %
    We have modified the example on \textbf{pages 16--17} to include more
    explanatory text and to use Python's built-in
    multiprocessing library to run the simulations on many processors. We note that if Cantera has
    been built with multi-threaded BLAS and LAPACK libraries (such as the Intel MKL), then the 0-D
    solver will automatically utilize the multithreading capabilities of those libraries. Therefore,
    using the multiprocessing library may not give much benefit in terms of overall speedup,
    depending on the complexity of the chemical kinetic model (i.e., how long the simulation
    takes to run), how many processors are available, and the availability of the multi-threaded
    BLAS and LAPACk libraries. However, we have not explained this detail of Cantera in the text to
    avoid confusing the reader.
    %
\end{response}

\subsection*{Conclusions and future work}

\begin{reviewer}
    %
    29.~No conclusion is mentioned for the usage examples.
    %
\end{reviewer}

\begin{response}
    %
    We have added a paragraph to the conclusion on \textbf{page 17}.
    %
\end{response}

### 2017-10-09 rebuttal.tex

\begin{reviewer}
    %
    p.~1 ``Competing standards such as the Cantera [2] CTI format or FlameMaster [3] lag behind
    considerably, although use of the former continually grows due to Cantera's open availability.''

    Note that FlameMaster is also freely available from the GitLab repository of the developers.
    %
\end{reviewer}

\begin{response}
    We attempted to access the GitLab repository at \url{https://git.rwth-aachen.de/ITV/FlameMaster}, but
    it is a closed repository. Thus, while it may be freely available to those with access, it is not
    \emph{openly} available.

    To clarify the meaning of ``open'' when used in our article, we have added to \textbf{page 2} a
    brief definition and link to the definition of open source, which comes from the
    Open Source Initiative and the Open Knowledge Institute.
\end{response}

@@

\begin{reviewer}
    %
    12.~Is it possible to create for example two \texttt{common-properties} \texttt{composition}
    blocks A and B and refer to them in the \texttt{DataPoint} keys?
    %
\end{reviewer}

\begin{response}
    %
    At present, this is not possible; only a single \verb|composition| block is allowed in the
    \verb|common-properties| section. However, we are considering modifying the behavior of the
    \verb|common-properties| section, and discussion is ongoing in
    \url{https://github.com/pr-omethe-us/PyKED/issues/67}.
    %
\end{response}

\begin{reviewer}
    %
    13.~Page 6, line 12 describes the initial efforts of building an open repository of ChemKED
    files, on GitHub. Is this the most efficient way for making a publically available database? One
    issue with this is the fact that adding files is done via pull requests, which have to be
    verified and approved by the owners of the repository. Therefore a certain delay is inevitable.
    %
\end{reviewer}

\begin{response}
    %
    We agree that submitting files to the ChemKED database on GitHub will inevitably incur some
    delay; however, we see the human intervention step as necessary and important. In much the same
    way that review of manuscripts provides a certain surety of the minimal quality of the article,
    we see review of data being added to the database as a minimal quality check; for instance, does
    the reference exist, has the data already been added to the database, are the files styled
    consistently, etc. Furthermore, such human review can help to catch some errors that may be
    introduced by manual creation of the files, as pointed out by the reviewer previously.

@@

    Second, the specific feature mentioned here will be something that the user explicitly requests (e.g., by
    attempting to convert mass to mole fractions). Thus, we expect that they are willing to wait for
    the connection to complete to use the feature. Third, we expect that the amount of data
    transmitted for such a request would be small, on the order of a few kB, so that even slow
    internet connections would be able to handle such requests. Finally, in addition to
    internet-based options, we are also investigating the use of standalone packages, such as RDKit,
    that would enable local retrieval of such information.
    %
\end{response}

\begin{reviewer}
    %
    22.~Page 7, line 40: Why is it interesting to create the YAML file again with the
    \texttt{write\_file()} function. Creating a summary excel/CSV file would be more useful.
    %
\end{reviewer}

\begin{response}
    %
    The use case for the \texttt{write\_file()} is that users might modify the ChemKED instance
    after it has been loaded, and wish to write their changes out to a new file.

    Users can create Excel or CSV files containing all or some of the datapoints in the ChemKED
    instance by using the \verb|get_dataframe()| function, and the associated functionality in
    Pandas (\verb|DataFrame.to_csv()| and \verb|DataFrame.to_excel()|). Since this functionality
    already exists, we do not wish to reinvent the wheel by writing our own functions that would
    have to be maintained and tested. Finally, we would note that data can also be plotted and
    viewed directly from Python using existing functionality in the Pandas DataFrame class
    (\verb|DataFrame.plot()|), which leverages the Matplotlib package under-the-hood.
    %
\end{response}

### 2017-10-09 rebuttal.tex

\setlength{\parindent}{0in}
\setlength{\parskip}{0.25cm}

\newenvironment{reviewer}{\vspace{0.5\baselineskip}\begingroup\itshape\textbf{Reviewer:}}{\endgroup}
\newenvironment{response}{\textbf{Our Response:}}{\vspace{0.5\baselineskip}}
\newcommand{\review}[1]{{\itshape#1}}
\newcommand{\respond}[1]{\textbf{Our Response:} #1}

\begin{document}

We thank the reviewers for their helpful and detailed comments, and have addressed their remarks
here. The manuscript has been significantly revised, and we have included a marked copy of the
manuscript with additions in blue and deletions in red.

### 2017-10-09 rebuttal.tex

\begin{response}
    %
    We have edited that sentence on \textbf{page 3} to correct the error.
    %
\end{response}

\subsection*{Overview of ChemKED format}

\begin{reviewer}
    %
    9.~A reference should not be a required mapping of an experimental dataset. If the intention is
    to make the community to adopt the ChemKED file structure, it might be easier to do so if
    experimentalists can use the structure internally as well, i.e. to read, store, and share
    non-published internal data.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that some of the choices currently enforced in the PyKED package and
    the ChemKED schema do not lend themselves to easy informal\slash working use. We presently recommend that
    users who wish to use ChemKED files in this way include an incomplete reference field. We have
    modified the ChemKED schema (and \textbf{page 5} of the manuscript) in response to this comment
    so that the only required subfields of the reference field are the authors and the year.

### 2017-10-09 pyked.tex
minor wording edits in conclusions

\ck{} and \pk{} are currently under active development, and contributions from the community are
welcomed. All development occurs under the BSD 3-clause open source license and the code is housed
on GitHub at \url{https://github.com/pr-omethe-us/PyKED}. Future directions for development are
outlined in the public roadmap.\footnote{\url{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}}
The highest priority issues are currently adding support for other types of fundamental experiments,
including speciation measurements in shock tubes, jet-stirred reactors, and flow reactors, and
laminar flame speed and flame extinction measurements. Questions, comments, or suggestions are
welcomed and can be posted as issues in the GitHub repository or emailed to the authors. In
particular, we welcome proposals for new experimental types; see issue \#60 in the \pk{} repository
for an example of how this may be
done.\footnote{\url{https://github.com/pr-omethe-us/PyKED/issues/60}}

>>>>

Finally, we presented several examples using \pk{} to interpret and interact with data
stored in \ck{} files. Both examples are also available online as Jupyter Notebook files in the
GitHub repository and in the Supplementary Material. The first example simulates a
rapid compression machine experiment, including the facility-specific effects critical to
accurately predict ignition delay experiments. The second example demonstrates the use of built-in
Python libraries to automatically simulate several shock tube experiments in parallel,
with initial conditions loaded directly from the \ck{} file.

We are actively developing \ck{} and \pk{}, and we welcome contributions from the community.
All development occurs under the BSD 3-clause open source license and the code is housed
on GitHub at \url{https://github.com/pr-omethe-us/PyKED}. Future directions for development are
outlined in the public roadmap.\footnote{\url{https://github.com/pr-omethe-us/PyKED/wiki/Roadmap}}
The highest priority issues are currently adding support for other types of fundamental experiments,
including speciation measurements in shock tubes, jet-stirred reactors, and flow reactors, and
laminar flame speed and flame extinction measurements. Questions, comments, or suggestions are
welcomed and can be posted as issues in the GitHub repository or emailed to the authors. In
particular, we welcome proposals for new experimental types; see issue \#60 in the \pk{} repository
for an example of how this may be
done.\footnote{\url{https://github.com/pr-omethe-us/PyKED/issues/60}}
